Gender, Age, and Peer Influence on Food Incentives in Schools

### 2024-02-20 outputs/paper/paper.qmd
updates

In the @tbl-table1, the results are generally consistent across the 4 subgroups, in which there is low variance among the values in the various columns. This more-or-less even distribution of the races, genders and grades of students across tables of varying incentive levels speaks to the efficacy of the randomness that was intended in the original study. In the next section, we will dive into analyzing the correlation of several individual factors and the choices made in stage G2.

@@

To begin our analysis, we created a graph, @fig-foodChoices, of the frequency of students that either switched their final food choice or kept the same food choice, grouped by the grade of the student. Grade 0 represents students in kindergarten. @fig-foodChoices gives an overview of the public data across all levels of table incentivises, in which students were able to make their final choice based on their peers' initial choices and incentive statuses.
\newpage

@@

This paper replicated and reproduced certain aspects of the paper *Incentives and Unintended Consequences: Spillover Effects in Food Choice* [@angelucci_incentives_2019]. @fig-lunchStatus and @fig-grade were replicated using the original raw data provided in the replication package as well as Table 1. We discussed the original study and how they conducted their research as well as their original conclusions.

The purpose of our paper was to analyze the influence that age, gender, and free lunch eligibility may have had on a student’s food choice in a public setting where other peers’ choices and incentive status were visible to all students at the table. Incentive status plays an important role in affecting each student’s food choice decision, as the possibility to win a prize is offered to an incentivized student when that student chooses grapes. The results of these individual analyses are mentioned in the Results section. To begin the discussion of our analyses, we will first look at the results for @fig-lunchStatus, in which the impact of free lunch eligibility on incentivized students’ food choices is analyzed.

@@

Speak about any holes in our analysis or perhaps the original study. One point worth mentioning may be the fact that the study was undertaken in a low-income neighborhood in Chicago. When introducing the incentive factor in the public treatment, the spillover effect may be magnified as these elementary students may be more influenced to alter their final food choice and choose the incentivized grapes, compared to students surveyed in more affluent neighborhoods who may not be as swayed by the option of a prize.

>>>>

In @tbl-table1, the results are generally consistent across the 4 subgroups, in which there is low variance among the values in the various columns. This more-or-less even distribution of the races, genders and grades of students across tables of varying incentive levels speaks to the efficacy of the randomness that was intended in the original study. In the next section, we will dive into analyzing the correlation of several individual factors and the choices made in stage G2.

@@

To begin our analysis, we created a graph, @fig-foodChoices, of the frequency of students that either switched their final food choice or kept the same food choice, grouped by the grade of the student. Grade 0 represents students in kindergarten. @fig-foodChoices gives an overview of the public data across all levels of table incentivization, in which students were able to make their final choice based on their peers' initial choices and incentive statuses.
\newpage

@@

This paper replicated and reproduced certain aspects of the paper *Incentives and Unintended Consequences: Spillover Effects in Food Choice* [@angelucci_incentives_2019]. @fig-lunchStatus and @fig-grade were replicated using the original raw data provided in the replication package as well as @fig-foodChoices. We discussed the original study and how they conducted their research as well as their original conclusions.

The purpose of our paper was to analyze the influence that age, gender, and free lunch eligibility may have had on a student’s food choice in a public setting where other peers’ choices and incentive status were visible to all students at the table. Incentive status plays an important role in affecting each student’s food choice decision, as the possibility to win a prize is offered to an incentivized student when that student chooses grapes. The results of these individual analyses are mentioned in @sec-results. To begin the discussion of our analyses, we will first consider the results for @fig-lunchStatus, in which the impact of free lunch eligibility on incentivized students’ food choices is analyzed.

@@

Being that the study was undertaken in a low-income neighborhood in Chicago, when introducing the incentive factor in the public treatment, the spillover effect may be magnified as these elementary students may be more influenced to alter their final food choice and choose the incentivized grapes. As this study took place in a low-income neighbourhood in Chicago, IL, it is rewasonable to say that this study was conducted in a socio-cultural vaccum. Compared to students surveyed in more affluent neighborhoods who may not be as swayed by the option of a prize, students in the state of Chicago qualify for free lunch because their annual household income is below the poverty line--the impact of this socio-economic experience cannot be captured in a study of such nature, nor a study of such size (diminutive compared to the scale of the impact of the lived experience). Students could have been making decisions based off of food scarcity, resource scarcity, peer pressure, fear of judgement, fear of authority, so on and so forth. So, while it is completely fair to judge the results of the original study for what they are, it is not ethical to make assumptions or inferences about *why* students make the decisions they make with regard to this study, without a comparison study or a larger data set to anchor statistics across demographics outside of Chicago, or outside of this specific school. Ultimately, it is unequivocally impossible to account for and measure every variable which influenced the students' decision-making, incentivized or not.

Regardless of the nature of the decision-making, it is clear that incentivization works. Looking toward the future, we may consider: how do we leverage incentivization to produce large-scale results and changes in behaviour? For instance, if the balh balh balh
